---
title: "DGRP Genetic Arch"
author: "Alexander Shingleton"
date: "9/8/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressPackageStartupMessages({
 require(lme4)
require(lmerTest)
require(lattice) # for some of the simple plots
require(MCMCglmm) # To get posterior distributions
library(pbkrtest) # parametric bootstrap.
require(smatr)
require(dplyr)
require(plyr)
require(rrcov)
require(tibble)
require(ggplot2)
require(ggalt)
require(varhandle)
require(magrittr)
require("ggthemes")
 })

```

Preamble


```{r,message=FALSE, warning=FALSE}

df <- read.csv("~/Documents/Data/DGRP Genetic Architecture/DGRPfinal_clean.csv")
summary(df)
df<-na.omit(df)
df$line<-as.factor(df$line)
df$day<-as.factor(df$day)
```
#Is there variation among blocks

Look at the control lineages to see whether there is variation among blocks
```{r,message=FALSE, warning=FALSE}
dfc<-subset(df0, type=="ct")
blocktest<-lmer(pupa~sex*block+(1|line), data=dfc)
Anova(blocktest, type="III")
```

#SSD

Is there SSD? Using only fed flies (day=0)

```{r,message=FALSE, warning=FALSE}
df0<-subset(df, day==0)
df0<-na.omit(df0)
SSD<-lmer(pupa~sex+(1|line) +(1|block), data=df0)
summary(SSD)
anova(SSD)
```
Does SSD vary among lineages?

```{r,message=FALSE, warning=FALSE}
model2<-lmer(pupa~sex+(1|line)+(1|block), data=df0)
model1<-lmer(pupa~sex+(sex|line)+(1|block), data=df0)
summary(model1)
anova(model1,model2)
```
Do a LRT

```{r,message=FALSE, warning=FALSE}
(AIC(model1) - REMLcrit(model1))/2 # # of parameters the model "thinks" are being estimated
(AIC(model2) - REMLcrit(model2))/2 # # of parameters the model "thinks" are being estimated

```
So lme4/lmer is treating model 1 as having two more parameters than model2.
```{r,message=FALSE, warning=FALSE}
LR.model <-  -as.numeric(REMLcrit(model1) - REMLcrit(model2))
LR.model
nlevels(df$line)
pchisq(q = LR.model, df=2, lower=F)
pchisq(q = LR.model, df=nlevels(df$line), lower=F)
```
Finally, we can conduct a parametric bootstrap to compare the two models.

```{r echo=FALSE, message = FALSE, warning=FALSE}
PBmodcomp(model1, model2, nsim=1000, details = 1)
summary(pbtest)
```

Finally using Bayesian Analysis

```{r,message=FALSE, warning=FALSE}
prior.2 <-list(R=list(V=0.01, nu=0.002), 
               G=list(G1=list(V=0.01*diag(1), nu=0.002),
                      G2=list(V=0.01*diag(2), nu=0.002)))

model1M.MCMC <- MCMCglmm(pupa ~ 1 + sex, 
  random=~block + us(1 + sex):line,
  prior = prior.2, burnin = 5000, nitt = 20000, thin = 10,
  verbose = F, pr = T,
  data=df0)
summary(model1M.MCMC)
```

#SSP

Does SSP vary among lineages?

```{r,message=FALSE, warning=FALSE}

model2<-lmer(pupa~sex*day+(sex+day|line)+(1|block), data=df)
model1<-lmer(pupa~sex*day+(sex*day|line)+(1|block), data=df)
anova(model1,model2)
```