---
title: "DGRP Genetic Arch"
author: "Alexander Shingleton"
date: "9/8/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressPackageStartupMessages({
 require(lme4)
require(lmerTest)
require(lattice) # for some of the simple plots
require(MCMCglmm) # To get posterior distributions
require(smatr)
require(dplyr)
require(plyr)
require(rrcov)
require(tibble)
require(ggplot2)
require(ggalt)
require(varhandle)
require(magrittr)
require("ggthemes")
require(pbkrtest)  # parametric bootstrap
library(car)
 })
R.version
```




#Preamble
```{r,message=FALSE, warning=FALSE}
#dataset was manually relabeled
df <- read.csv("~/Dropbox/_Github_reps/DGRP_SSDSSP/Sept2020/DGRPfinal_clean_labels.csv")

#line and day of treatment as categorical
df$line<-as.factor(df$line)
df$day<-as.factor(df$day)
df$block<-as.factor(df$block)
#summary of data
summary(df)

#remove all NAs, most functions do not deal with NAs well
df<-na.omit(df)


```



# Filtering out lines with few data
```{r}
#subsetting experimental flies only

df_exp<-subset(df, type == "exp")
df<-df_exp

#filtering out groups (line x sex x day) that have less than 10 flies
df_f<-subset(df, sex =="F")
df_f0<-subset(df_f, day =="0")
df_f1<-subset(df_f, day =="1")
df_f2<-subset(df_f, day =="2")

#F0
df_subf0<-df_f0%>%
group_by(line) %>%
filter(n() >=10)

length(unique(df_subf0$line)) 

#F1
df_subf1<-df_f1%>%
group_by(line) %>%
filter(n() >=10)

length(unique(df_subf1$line))  

#F2
df_subf2<-df_f2%>%
group_by(line) %>%
filter(n() >=10)

length(unique(df_subf2$line))  

df_m<-subset(df, sex =="M")
df_m0<-subset(df_m, day =="0")
df_m1<-subset(df_m, day =="1")
df_m2<-subset(df_m, day =="2")

#M0
df_subm0<-df_m0%>%
group_by(line) %>%
filter(n() >=10)

length(unique(df_subm0$line)) 

#M1
df_subm1<-df_m1%>%
group_by(line) %>%
filter(n() >=10)

length(unique(df_subm1$line))  

#M2
df_subm2<-df_m2%>%
group_by(line) %>%
filter(n() >=10)

length(unique(df_subm2$line))  



```

#rebinding dataframes

```{r}
df0<-rbind(df_subm0, df_subf0)
df0<-na.omit(df0)
length(unique(df0$line))  #182 lineages
length(unique(df0$type))

```


#Is there variation among blocks
Flies of lineages were collected in different blocks, periods of time. Control lineages were selected and collected repeatedly for each block so that if there is variation, we can account for that factor.

Look at the control lineages to see whether there is variation among blocks

```{r,message=FALSE, warning=FALSE}
#subset control lineages
df0<-subset(df, day =='0')
dfc<-subset(df0, type=="ct")
dfcf<-subset(dfc, sex=="F")
dfcm<-subset(dfc, sex=="M")


#linear mixed model testing pupa size variation and if block factor has an effect on the variation
blocktest<-lmer(pupa~block+(1|line), data=dfc)
Anova(blocktest, type="III")

blocktest2<-lmer(pupa~sex*block+(1|line), data=dfc)
Anova(blocktest2, type="III")

blocktestF<-lmer(pupa~block+(1|line), data=dfcf)
Anova(blocktest, type="III")

blocktestM<-lmer(pupa~block+(1|line), data=dfcm)
Anova(blocktest, type="III")
```

There is a difference between collecting blocks so we have to account for block as a random factor, or adjust the values so there is no variation between blocks.




# SSD

## Is there SSD? Using only fed flies (day=0)
To test if we have sexual size dimorphism, we want to test the effect of size with line and block as random factor. If there is a variation due to sex, that means SSD is present.

```{r,message=FALSE, warning=FALSE}
#subsetting day 0, fed flies
df0<-subset(df, day==0)

SSD<-lmer(pupa~sex+(1|line) +(1|block), REML=TRUE, data=df0)
summary(SSD)
anova(SSD)
```

## Does SSD vary among lineages?
Now we want to know if that SSD varies between lines.

```{r,message=FALSE, warning=FALSE}

model2<-lmer(pupa~sex+(1|line)+(1|block), data=df0)  #model to test for SSD presence as we did above
model1<-lmer(pupa~sex+(sex|line)+(1|block), data=df0) 
anova(model1)
anova(model1,model2)
```
Model 1 is better, as AIC and BIC is smaller and log likelihood is higher. The difference of fit between these two models is significant.

Do a LRT

```{r,message=FALSE, warning=FALSE}
(AIC(model1) - REMLcrit(model1))/2 # # of parameters the model "thinks" are being estimated
(AIC(model2) - REMLcrit(model2))/2 # # of parameters the model "thinks" are being estimated

```


So lme4/lmer is treating model 1 as having two more parameters than model2.
```{r,message=FALSE, warning=FALSE}
LR.model <-  -as.numeric(REMLcrit(model1) - REMLcrit(model2))
LR.model
nlevels(df$line)
pchisq(q = LR.model, df=2, lower=F)
pchisq(q = LR.model, df=nlevels(df$line), lower=F)
```
Finally, we can conduct a parametric bootstrap to compare the two models.

```{r echo=FALSE, message = FALSE, warning=FALSE}
pbtest<-PBmodcomp(model1, model2, nsim=1000, details = 1)
summary(pbtest)
```

Finally using Bayesian Analysis

```{r,message=FALSE, warning=FALSE}
prior.2 <-list(R=list(V=0.01, nu=0.002), 
               G=list(G1=list(V=0.01*diag(1), nu=0.002),
                      G2=list(V=0.01*diag(2), nu=0.002)))

model1M.MCMC <- MCMCglmm(pupa ~ 1 + sex, 
  random=~block + us(1 + sex):line,
  prior = prior.2, burnin = 5000, nitt = 20000, thin = 10,
  verbose = F, pr = T,
  data=df0)
summary(model1M.MCMC)
```



### Post model fitting check

#### Residual distribution
```{r,message=FALSE, warning=FALSE}
res_model1=residuals(model1)
```
Model 1 residual distribution

```{r,message=FALSE, warning=FALSE}
plot(model1)
```



Model 1 residual distribution per random factor


QQ plot
```{r,message=FALSE, warning=FALSE}
qqnorm(res_model1)
```


Random effect plot
```{r,message=FALSE, warning=FALSE}
qqmath(ranef(model1))
```
#SSP

Does SSP vary among lineages?

```{r,message=FALSE, warning=FALSE}
# we want to compare effect of sex and day of starvation on pupal size.
model2<-lmer(pupa~sex*day+(sex+day|line)+(1|block), data=df)
model1<-lmer(pupa~sex*day+(sex*day|line)+(1|block), data=df)
anova(model1,model2)
```


#Plasticity

Do we have variation in plasticity in females?
```{r,message=FALSE, warning=FALSE}
dfcF<-subset(df, df$type == "ct" & df$sex =="F")

dfcM<-subset(df, df$type == "ct" & df$sex =="M")


model1F<-lmer(pupa~day+(1|line)+(1|block), data=dfcF)
model2F<-lmer(pupa~day+(day|line)+(1|block), data=dfcF)
anova(model1F, model2F)



```


Do we have variation fo plasticity in males?






#WINGS

### How much bigger are female wings compared to males

```{r,message=FALSE, warning=FALSE}
df0$sex

wingdf0<-subset(df0, df0$wing >=12)
meanwing<-tapply(df0$wing, df0$sex, summary)

(exp(meanwing$F[[4]]) -exp(meanwing$M[[4]]))/exp(meanwing$M[[4]])*100

```
Wings are 21% larger in females than males
