---
title: "DGRP Genetic Arch"
author: "Alexander Shingleton"
date: "9/8/2020"
output:
  word_document: default
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressPackageStartupMessages({
 require(lme4)
require(lmerTest)
require(lattice) # for some of the simple plots
require(MCMCglmm) # To get posterior distributions
require(smatr)
require(dplyr)
require(plyr)
require(rrcov)
require(tibble)
require(ggplot2)
require(ggalt)
require(varhandle)
require(magrittr)
require("ggthemes")
require(pbkrtest)  # parametric bootstrap
require(car)
require(tidyverse)
require(hrbrthemes)
require(viridis)
require(tidyr)
require(PerformanceAnalytics)
require(psych)
require(Deducer)
require(ggpubr)
 })
R.version
```
# Sept 29 2020
 - aggregate plots to figures by section

#Pre-processing of data
  - n<10 groups removed
  - remove block effect when calculating SSD and SSP 

# Outline of questions

- Section 1: SSD in fed flies
  - Question 1.1: Is there Sexual Size Dimorphism in the DGRP lines.
  - Question 1.2: Do we see a genetic variation of SSD in the DGRP lines.
  - Question 1.3: Is the SSD genetic variation due to a genetic variation in male or in female size?

- Section 2: SSD in starved flies - use D1
  - Question 2.1: Is SSD in starved flies the same as in fed flies? (do a lmer sex*day on D0 and D1)
    -  Does SSD increase or decrease when the flies are starved? Our hypothesis is that overall SSD should decrease
  - Question 2.2: Does the SSD in starved conditions vary the same way as in fed flies?
    - use a correlation and variance analysis

We have established that SSD0 and SSD1 are not the same. If they were the same, that would mean that the body size of male and female changed in the same way in different environment. We know that males and females do not change the same way so that means we have Sex Specific Plasticity. 

-Section 3: Male and Female Plasticity
  - Question 3.1: Which sex is more plastic and does plasticity vary among lines?
  - Question 3.2: Which sex plasticity contributes to SSP.
  
# Data preparation
importing file and checking for column names
```{r,message=FALSE, warning=FALSE}
#dataset was manually relabeled and control lines removed
df <- read.csv("~/Dropbox/_Github_reps/DGRP_SSDSSP/Data/DGRPfinal_clean_first_ctrl_only.csv")

#column names
names(df)
```

add a letter in front of factorial columns
```{r}
#add character in front of line, day and block
df$line <- sprintf('L%i', df$line)
df$day<-sprintf('D%i', df$day)
df$block<-sprintf('B%i', df$block)
#summary of data
summary(df)
```

## Filtering out lines with few data
```{r}
#first, create a column to indicate group by line_sex_day. This will be easier to filter out or subset by group later
df$group<- paste(df$line, df$sex, df$day, sep = "_")
head(df,2)
```

### KEEP THIS HERE BUT RUN BEFORE ANALYSES
```{r}
#filtering out groups (line x sex x day) that have less than 10 flies
#NB: check if I run that filter after na omit before each analysis might not be better?
#df_sub<-df%>%
#group_by(group) %>%
#filter(n() >=10)
```



## Is there variation among blocks = OK
Flies of lineages were collected in different blocks, periods of time. Control lineages were selected and collected repeatedly for each block so that if there is variation, we can account for that factor.

Look at the control lineages to see whether there is variation among blocks

```{r,message=FALSE, warning=FALSE}
#removing all groups with less than 10 flies
df_sub<-df%>%
group_by(group) %>%
filter(n() >=10)  

#how many Na in pupae
sum(is.na(df_sub$pupa)) #1970 Nas

#remove NA in wing column
df_sub<-df_sub[!is.na(df_sub$pupa), ]

#subset fed and 1day starved flies
df_sub01<-subset(df_sub, day=="D0"|day=="D1")

#lm with block fixed effect
blocktest01<-lm(pupa~block, data=df_sub01)
summary(blocktest01)
Anova(blocktest01)

#move residuals to dataframe
df_sub01$pupa_noblock <- blocktest01$resid
head(df_sub01)

#df_sub01$pupanoblock is the column I will use when I want to calculate summary and indexes for correlation analyses

#rerun lm on pupa_noblock
blocktest01_2<-lm(pupa_noblock~block, data=df_sub01)
summary(blocktest01_2)
Anova(blocktest01_2) #block effect is gone!

```

There is a difference between collecting blocks so we have to account for block as a random factor when using lmer, or use the adjusted values to calculate SSD and SSP values.


# Section 1: SSD in fed flies
  - Question 1.1: Is there Sexual Size Dimorphism in the DGRP lines. A1.1: YES
  Plot box plot -> should I use adjusted value?
  - Question 1.2: Do we see a genetic variation of SSD in the DGRP lines. A1.2: YES
  Plot mean female size and male size on a bar graph per line
  Plot SSD on a separate bar graph
  
  - Question 1.3: Is the SSD genetic variation due to a genetic variation in male or in female size? FEMALE
  Plot correlation between SSD0 and female size, and SSD0 and male size in same Figure


## Question 1.1: Is there Sexual Size Dimorphism in the DGRP flies that are normally fed?
To test if we have sexual size dimorphism, we want to test the effect of size with line and block as random factor. If there is a variation due to sex, that means SSD is present.

```{r,message=FALSE, warning=FALSE}
df<-df_sub01  #do not forget to reload original csv data if I want to filter out things differently

#subsetting day 0, fed flies
df0<-subset(df, day=="D0") #8500
head(df0,2) 

#na.omit only if pupa has NA if not done before
#df0<-na.omit(df0, cols="pupa") 
```

### Linear Mixed Model on fed flies
```{r}
# Testing effect of sex in pupa size, with random effect for line and block.
SSDtest<-lmer(pupa~sex+(1|line) +(1|block), REML=TRUE, data=df0)
summary(SSDtest) #Number of obs: 8065, groups:  line, 187; block, 9
Anova(SSDtest)
plot(SSDtest)
qqnorm(resid(SSDtest))
qqline(resid(SSDtest))
```
### Box plot showing average difference in pupal size between male and female => ok
Plot with block effect
```{r}
# original pupa size
(SSDtest_plot <- ggplot(df0, aes(x = sex, y = pupa, colour = sex)) +
      geom_boxplot(alpha=0.3) +
      theme_classic() +
   geom_jitter(color="black", size=0.4, alpha=0.2)+
      theme(legend.position = "none",
            panel.spacing = unit(1, "lines")) +  # adding space between panels
   ggtitle("Body Sexual Size Dimorphism in DGRP flies")
 )

```
Plot adjusted no block effect
```{r}
#removed block effect pupal size

(SSDtest_plot_noblock <- ggplot(df0, aes(x = sex, y = pupa_noblock, colour = sex)) +
      geom_boxplot(alpha=0.3) +
      theme_classic() +
   geom_jitter(color="black", size=0.4, alpha=0.2)+
 
      theme(legend.position = "none",
            panel.spacing = unit(1, "lines")) +  # adding space between panels
   ggtitle("Body Sexual Size Dimorphism in DGRP flies no block effect")
 )
```

Answer 1.1: Yes, there is sexual size dimorphism in the DGRP flies, and females are larger than males on average.


## Question 1.2: Do we see a genetic variation of SSD in the DGRP lines?
### Comparing two models using ANOVA
```{r,message=FALSE, warning=FALSE}
#comparing two models
model2<-lmer(pupa~sex+(1|line)+(1|block), data=df0)  #model to test for SSD presence as we did above
model1<-lmer(pupa~sex+(sex|line)+(1|block), data=df0) 
anova(model1)
anova(model1,model2)
```
Model 1 is better, as AIC and BIC is smaller and log likelihood is higher. The difference of fit between these two models is significant.


```{r}
summary(model1)
```
In the model summary we see that the fitted value is smaller for males, so SSD is female biased even when considering genetic variation of SSD.

#### Do a LRT
How many parameters for each models
```{r,message=FALSE, warning=FALSE}
(AIC(model1) - REMLcrit(model1))/2 # # of parameters the model "thinks" are being estimated
(AIC(model2) - REMLcrit(model2))/2 # # of parameters the model "thinks" are being estimated

```


So lme4/lmer is treating model 1 as having two more parameters than model2.
```{r,message=FALSE, warning=FALSE}
LR.model <-  -as.numeric(REMLcrit(model1) - REMLcrit(model2))
LR.model
nlevels(df$line)
pchisq(q = LR.model, df=2, lower=F)
pchisq(q = LR.model, df=nlevels(df$line), lower=F)
```

#### Parametric boostrap
Finally, we can conduct a parametric bootstrap to compare the two models.

```{r echo=FALSE, message = FALSE, warning=FALSE}
#too long, will run later
#pbtest<-PBmodcomp(model1, model2, nsim=1000, details = 1)
#summary(pbtest)
```

#### Finally using Bayesian Analysis

```{r,message=FALSE, warning=FALSE}
#prior.2 <-list(R=list(V=0.01, nu=0.002), 
 #              G=list(G1=list(V=0.01*diag(1), nu=0.002),
  #                    G2=list(V=0.01*diag(2), nu=0.002)))

#model1M.MCMC <- MCMCglmm(pupa ~ 1 + sex, 
 # random=~block + us(1 + sex):line,
  #prior = prior.2, burnin = 5000, nitt = 20000, thin = 10,
  #verbose = F, pr = T,
  #data=df0)
#summary(model1M.MCMC)
```



#### Post model 1 fitting check

##### Residual distribution
```{r,message=FALSE, warning=FALSE}
```

##### Model 1 residual distribution

```{r,message=FALSE, warning=FALSE}
plot(model1)
```




##### QQ plot
```{r,message=FALSE, warning=FALSE}

res_model1=residuals(model1)
ggqqplot(res_model1)
```


##### Random effect plot
```{r,message=FALSE, warning=FALSE}
qqmath(ranef(model1))
```


### Plot mean size female and male per line + SSD per line
Calculate mean size per line
```{r}
#calculate means for each group using pupa_noblock
head(df0,2)
df0_mean<-aggregate(df0[, 10], list(df0$group), mean)
head(df0_mean,2)

#re-add line, day and sex columns
df0_mean<-df0_mean %>%
  separate(Group.1, c("line", "sex","day"), "_")
head(df0_mean,2)
```

Plot Mean size by line for fed flies
```{r}
meanpupa0_plot<-ggplot(df0_mean, aes(x=reorder(line,pupa_noblock), y=pupa_noblock)) +
  geom_col(aes(fill = sex)) +
  #facet_wrap(~ day) +
  #ggtitle("Mean pupa size per sex for each line in fed flies, without block effect")
theme(legend.direction = "horizontal",legend.position = "top",axis.title.x = element_blank(),axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
meanpupa0_plot
```

Calculate SSD index 
```{r}
#calculate SSD0
#separating males and females to put the values in columns
df0_mean_F<-subset(df0_mean, sex=="F")
df0_mean_M<-subset(df0_mean, sex=="M")

df0_mean_2<-merge(x=df0_mean_F, y=df0_mean_M, by.x="line", by.y="line")
head(df0_mean_2)

#remove extra columns
df0_mean_2<-df0_mean_2[,c(1,4,7)] #line, pupa_noblockF and pupa_noblockM
colnames(df0_mean_2) <- c("line", "pupaF", "pupaM") #rename col

SSD0<-df0_mean_2 #move back to df0_mean
SSD0$SSD<- SSD0$pupaF - SSD0$pupaM  #since we established that females are larger than males in general, SSD is female-male sizes
head(SSD0)#182 lines
```

```{r}
SSD0_plot<-ggplot(SSD0, aes(x=reorder(line,pupaF), y=SSD)) +
  geom_col(aes()) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  labs(x = "DGRP line")

SSD0_plot

```

#### Arrange plots together
```{r}
ggarrange(meanpupa0_plot, SSD0_plot, 
          labels = c("A", "B"),
          ncol = 1, nrow = 2)
```


## Question 1.3: Is SSD genetic variation due to the genetic variation in male or in female size?

Correlation test and plot
```{r}
#reshape data
SSD0_gather<-gather(SSD0,key="sex",value="pupa",pupaF:pupaM)
#plot
sp <- ggscatter(SSD0_gather, x = "SSD", y = "pupa",
   color = "sex",
   add = "reg.line", conf.int = TRUE, alpha=0.5)
sp + ggpubr::stat_cor(aes(color = sex)) #stat_cor uses the Pearson correlation
```

In fed flies, SSD covaries with female size.





# Section 2: SSD in starved flies
We found that in fed flies, SSD is female biased and it varies genetically. We finally found out that SSD covaries with female size.

How is SSD in starved conditions?

- Section 2: SSD in starved flies - use D1
  - Question 2.1: Is SSD in starved flies the same as in fed flies? (do a lmer sex*day on D0 and D1)
    -  Does SSD increase or decrease when the flies are starved? Our hypothesis is that overall SSD should decrease
  - Question 2.2: Does the SSD in starved conditions vary the same way as in fed flies?
    - use a correlation and variance analysis
  
## Question 2.1: Do we have the same SSD when we change environment?
### 1 day starvation data preparation = OK
```{r}
#fisrt subset Day1 data from df_sub
df1<-subset(df_sub01, day=="D1") 
#na.omit only if pupa has NA
df1<-df1[!is.na(df1$pupa), ] #just in case, but already removed previously
df1<-df1%>%
group_by(group) %>%
filter(n() >=10)  
length(unique(df1$line)) #174 lines left
```

### Model fit for SSD
```{r}
# Testing effect of sex in pupa size, with random effect for line and block.
SSD1test<-lmer(pupa~sex+(1|line) +(1|block), REML=TRUE, data=df1)
summary(SSD1test)
Anova(SSD1test)
```
SSD still exists at 1 day starvation


### Plots 
Plot with block effect
```{r}
# original pupa size
(SSDtest_plot <- ggplot(df1, aes(x = sex, y = pupa, colour = sex)) +
      geom_boxplot(alpha=0.3) +
      theme_classic() +
   geom_jitter(color="black", size=0.4, alpha=0.2)+
      theme(legend.position = "none",
            panel.spacing = unit(1, "lines")) +  # adding space between panels
   ggtitle("Body Sexual Size Dimorphism in starved flies")
 )

```
Plot adjusted no block effect
```{r}
#removed block effect pupal size
(SSDtest_plot_noblock <- ggplot(df1, aes(x = sex, y = pupa_noblock, colour = sex)) +
      geom_boxplot(alpha=0.3) +
      theme_classic() +
   geom_jitter(color="black", size=0.4, alpha=0.2)+
 
      theme(legend.position = "none",
            panel.spacing = unit(1, "lines")) +  # adding space between panels
   ggtitle("Body Sexual Size Dimorphism in starved flies no block effect")
 )
```


HERE plot of mean size male and female for starved flies
```{r}
#calculate means for each group using pupa_noblock
df1_mean<-aggregate(df1[, 10], list(df1$group), mean)
head(df1_mean,2)

#re-add line, day and sex columns
df1_mean<-df1_mean %>%
  separate(Group.1, c("line", "sex","day"), "_")
head(df1_mean,2)
```

```{r}
meanpupa1_plot<-ggplot(df1_mean, aes(x=reorder(line,pupa_noblock), y=pupa_noblock)) +
  geom_col(aes(fill = sex)) +
  #facet_wrap(~ day) +
  #ggtitle("Mean pupa size per sex for each line in fed flies, without block effect")
theme(legend.direction = "horizontal",legend.position = "top",axis.title.x = element_blank(),axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),plot.title = element_text(size = 10, face = "bold"))
meanpupa1_plot+ggtitle("Mean pupa size (block effect removed) per DGRP line when starved, for each sex", )
```


##Question 2.2: Does overall SSD increase or decrease when the flies are starved? Our hypothesis is that overall SSD should decrease

Use df_sub01
```{r}
model2<-lmer(pupa~sex*day+(sex+day|line)+(1|block), data=df_sub01)
model1<-lmer(pupa~sex*day+(sex*day|line)+(1|block), data=df_sub01)
anova(model1,model2)
summary(model1)

```
Using summary model 1, I can calculate the means for Female 0, Male 0, Female 1 and Male 1.
```{r}
# figure out how to extract intercept values
F0 = 14.566230
M0 = 14.566230 -0.094440 
F1 = 14.566230 -0.179327
M1 = 14.566230 -0.094440 -0.179327 + 0.030618 
print("SSD0")
F0-M0
print("SSD1")
F1-M1
```
SSD1<SSD0 and it is significantly different

##### Calculating SSD in both conditions
```{r}

#separating males and females to put the values in columns
df1_mean_F<-subset(df1_mean, sex=="F")
df1_mean_M<-subset(df1_mean, sex=="M")

df1_mean_2<-merge(x=df1_mean_F, y=df1_mean_M, by.x="line", by.y="line")
head(df1_mean_2)

#remove extra columns
df1_mean_2<-df1_mean_2[,c(1,4,7)] #using pupa_noblock
colnames(df1_mean_2) <- c("line", "pupaF1", "pupaM1")

df1_mean<-df1_mean_2
df1_mean$SSD1<- df1_mean$pupaF1 - df1_mean$pupaM1  #since we established that females are larger than males in general
head(df1_mean) 
length(unique(df1_mean$line)) #157 lines left after calculating SSD1

#combining SSDs values without discarding rows, which would happen if I merged the dataframes (we would end up with 59 lines for all SSD2)
head(SSD0)
SSD0<-SSD0[,c(1,4)]
SSD0$day<-"D0"
head(SSD0)

SSD1<-df1_mean[,c(1,4)]
SSD1$day<-"D1"
names(SSD1)[names(SSD1) == "SSD1"] <- "SSD"

SSD_all<-rbind(SSD0,SSD1)
head(SSD_all)
```

SSD boxplot, the overall SSD decreases when flies are starved
```{r}
# plotting male and female size mean
SSD_all %>%
  ggplot( aes(x=day, y=SSD)) +
    geom_boxplot() +
    scale_fill_viridis(discrete = TRUE, alpha=0.6) +
    geom_jitter(color="black", size=0.4, alpha=0.2) +
    #theme_ipsum() +
    theme(
      legend.position="none",
      plot.title = element_text(size=11)
    ) +
    ggtitle("Sexual Size Dimorphism in DGRP flies at different starving conditions") +
    xlab("")

```
SSD decreases at Day1 starvation 


## Question 2.3: Does SSD in starved conditions vary, and does it covary with SSD0?
First, we want to see if there is genetic variation in SSD1 and SSD2
### Model fit
```{r}
#Comparing two model fit for SSD1 to see if there is genetic variation

model2<-lmer(pupa~sex+(1|line)+(1|block), data=df1)  #model to test for SSD presence as we did above
model1<-lmer(pupa~sex+(sex|line)+(1|block), data=df1) 
anova(model1)
anova(model1,model2)
```

Model 1 is better, as AIC and BIC is smaller and log likelihood is higher. The difference of fit between these two models is significant.

#### Do a LRT
How many parameters for each models
```{r,message=FALSE, warning=FALSE}
(AIC(model1) - REMLcrit(model1))/2 # # of parameters the model "thinks" are being estimated
(AIC(model2) - REMLcrit(model2))/2 # # of parameters the model "thinks" are being estimated

```


So lme4/lmer is treating model 1 as having two more parameters than model2.
```{r,message=FALSE, warning=FALSE}
LR.model <-  -as.numeric(REMLcrit(model1) - REMLcrit(model2))
LR.model
nlevels(df1$line)
pchisq(q = LR.model, df=2, lower=F)
pchisq(q = LR.model, df=nlevels(df1$line), lower=F)
```

#### Parametric boostrap
Finally, we can conduct a parametric bootstrap to compare the two models.

```{r echo=FALSE, message = FALSE, warning=FALSE}
#pbtest<-PBmodcomp(model1, model2, nsim=1000, details = 1)
#summary(pbtest)
```

#### Finally using Bayesian Analysis

```{r,message=FALSE, warning=FALSE}
#prior.2 <-list(R=list(V=0.01, nu=0.002), 
#               G=list(G1=list(V=0.01*diag(1), nu=0.002),
 #                     G2=list(V=0.01*diag(2), nu=0.002)))

#model1M.MCMC <- MCMCglmm(pupa ~ 1 + sex, 
#  random=~block + us(1 + sex):line,
 # prior = prior.2, burnin = 5000, nitt = 20000, thin = 10,
#  verbose = F, pr = T,
 # data=df1)
#summary(model1M.MCMC)
```

#### Post model 1 fitting check

##### Model 1 residual distribution

```{r,message=FALSE, warning=FALSE}
plot(model1)
```

##### QQ plot
```{r,message=FALSE, warning=FALSE}
res_model1=residuals(model1)
ggqqplot(res_model1)
```

##### Random effect plot
```{r,message=FALSE, warning=FALSE}
qqmath(ranef(model1))
```

SSD in starved flies varies genetically.

### Plot SSD in starved flies by line
```{r}
SSD1_plot<-ggplot(SSD1, aes(x=reorder(line,SSD), y=SSD)) +
  geom_col(aes()) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  labs(x = " DGRP line ",y="SSD starved flies")
SSD1_plot
```


### Covariation with SSD in fed flies?
Does SSD in fed flies covary with SSD in old flies?

To know, this I am looking a the correlation between SSD values in the two food conditions, across lines.
Plot of SSD values between two food conditions
```{r}
head(SSD_all,2)
ggplot(SSD_all, aes(x=reorder(line,SSD), y=SSD)) +
  geom_col(aes(fill = day)) +
  facet_wrap(~ day) +
  coord_flip()
```
Correlation between SSDs. Hypothesis: if SSD changes with environment, which is what we expect, we will not see a correlation between SSD0 and SSD1 and/or SSD2

```{r}
#reshape data
head(SSD_all)

x= SSD_all[SSD_all$day =="D0", ]
y= SSD_all[SSD_all$day =="D1", ]

SSD_merged<-merge(x=SSD_all[SSD_all$day =="D0", ], y=SSD_all[SSD_all$day =="D1", ], by.x="line", by.y="line") 
#plot
sp2<- ggscatter(SSD_merged, x="SSD.x" ,y="SSD.y",
   add = "reg.line", conf.int = TRUE, alpha=0.5)
sp2 + ggpubr::stat_cor(aes())#stat_cor uses the Pearson correlation

```
Correlation between SSD1 and SSD0, baarely significant and rsquare very low.

It seems that SSD1 may vary differently, does female covariation with SSD0 changes in starved conditions?

Correlation SSD1 and female and male size
```{r}
SSD1_gather<-gather(df1_mean, key="sex",value="pupa",pupaF1:pupaM1)
head(df1_mean) 
head(SSD1_gather)
sp3<- ggscatter(SSD1_gather, x="SSD1" ,y="pupa",
   color = "sex",
   add = "reg.line", conf.int = TRUE, alpha=0.5)
sp3 + ggpubr::stat_cor(aes(color=sex))#stat_cor uses the Pearson correlation

```
SSD1 also covaries with female size in starved conditions

Will still check SSD1, even though it does not correlate with SSD0, we can see if SSD1 variation covaries with either female or male variation

We have shown that SSD changes in different condition, that in starved conditions, we have a decrease in overall SSD, and that the variation of SSD partly covaries with female size. But the degree of correlation varies? (how to test that?)


SSD changes in different food conditions, that means that there is a specific plasticity between sexes, or Sex Specific Plasticity. How does it look like and which sex is responsible for this difference in plasticity that is linked to SSD?

# Section 3: SSP and Male, Female Plasticity
  - Question 3.1: confirming that SSP exists - same test as SSD0 and SSD1 comparison
  
  - Question 3.2: Does SSP vary genetically?
  
  - Question 3.2: Which sex is more plastic and does plasticity vary among lines?
  
  - Question 3.2: Which sex plasticity contributes to SSP.


  - Question 3.3: Is the variation in SSP due to a variation in SSD in fed or starved flies?
If SSP covaries with SSD0, that means that we have variation in fed flies. If SSP covaries with starved flies, we have variation in the reduction of flies size.
  - Question 3.4: If it is due to SSD in either fed or starved flies, is it because the females vary more or the males vary more in size?
  - Question 3.5: Does the variation causing SSP (Question 3.4) is the same as the variation caused by SSD in Question 1.3?


##Question 3.1: Confirming that SSP exists 
### SSP: between fed and 1 day starved flies
We showed in Question 2.X that SSD changed when changing food conditions. Let's view again the summary of the Model.

```{r}
model1<-lmer(pupa~sex*day+(sex*day|line)+(1|block), data=df_sub01)
summary(model1)
```

## Question 3.2: Does SSP vary in the DGRP flies?
### Model fit comparison - previously done to compare SSD0 and SSD1
```{r,message=FALSE, warning=FALSE}
# we want to compare effect of sex and day of starvation on pupal size. I am using df01, subset off data day 0 and day1

model2<-lmer(pupa~sex*day+(sex+day|line)+(1|block), data=df_sub01) #there is randomness of sex and also the effect of day
model1<-lmer(pupa~sex*day+(sex*day|line)+(1|block), data=df_sub01) #basic model
anova(model1,model2)
Anova(model1) #there is SSP
```

#### Do a LRT
How many parameters for each models
```{r,message=FALSE, warning=FALSE}
(AIC(model1) - REMLcrit(model1))/2 # # of parameters the model "thinks" are being estimated
(AIC(model2) - REMLcrit(model2))/2 # # of parameters the model "thinks" are being estimated

```


So lme4/lmer is treating model 1 as having five more parameters than model2.
```{r,message=FALSE, warning=FALSE}
LR.model <-  -as.numeric(REMLcrit(model1) - REMLcrit(model2))
LR.model
nlevels(df_sub01$line)
pchisq(q = LR.model, df=5, lower=F)
pchisq(q = LR.model, df=nlevels(df_sub01$line), lower=F)
```

#### Parametric boostrap
Finally, we can conduct a parametric bootstrap to compare the two models.

```{r echo=FALSE, message = FALSE, warning=FALSE}
#pbtest<-PBmodcomp(model1, model2, nsim=1000, details = 1)
#summary(pbtest)
```

#### Finally using Bayesian Analysis
#### NB: did not run before need to change the model
```{r,message=FALSE, warning=FALSE}
#prior.2 <-list(R=list(V=0.01, nu=0.002), 
  #             G=list(G1=list(V=0.01*diag(1), nu=0.002),
   #                   G2=list(V=0.01*diag(2), nu=0.002)))

#model1M.MCMC <- MCMCglmm(pupa ~ 1 + sex, #is this right for SSP?
  #random=~block + us(1 + sex):line,
  #prior = prior.2, burnin = 5000, nitt = 20000, thin = 10,
  #verbose = F, pr = T,
  #data=df01)
#summary(model1M.MCMC)
```



#### Post model 1 fitting check

#### Residual distribution
```{r,message=FALSE, warning=FALSE}
res_model1=residuals(model1)
```

#### Model 1 residual distribution

```{r,message=FALSE, warning=FALSE}
plot(model1)
```


#### QQ plot
```{r,message=FALSE, warning=FALSE}
require(ggpubr)
ggqqplot(res_model1)
```

#### Random effect plot
```{r,message=FALSE, warning=FALSE}
qqmath(ranef(model1))
```

Answer 3.2: There is genetic variation in SSP if we compare fed flies and starved flies.

### Plot of SSP per line
#### Calculate SSP
To calculate SSP, I first calculate the plasticity of female and male and subtract the plasticity female-male
```{r}
#calculate plasticity female

df1F<-subset(df1, sex=="F")
df1F_mean<-aggregate(df1F[, 10], list(df1F$line), mean) #using pupa_noblock
colnames(df1F_mean)<-c("line","pupaFmean_1")

df0F<-subset(df0, sex=="F")
df0F_mean<-aggregate(df0F[, 10], list(df0F$line), mean)
colnames(df0F_mean)<-c("line","pupaFmean_0")

pupa_mean_F01<-merge(x=df0F_mean, y=df1F_mean, by.x="line", by.y="line") 
head(pupa_mean_F01,2) #140 lines
pupa_mean_F01$PF01<-pupa_mean_F01$pupaFmean_0-pupa_mean_F01$pupaFmean_1 #calculating plast female 01
plastF01<-pupa_mean_F01 
```

```{r}
#calculate plasticity male
head(df1)
df1M<-subset(df1, sex=="M")
df1M_mean<-aggregate(df1M[, 10], list(df1M$line), mean)
colnames(df1M_mean)<-c("line","pupaMmean_1")
head(df1M_mean)

df0M<-subset(df0, sex=="M")
df0M_mean<-aggregate(df0M[, 10], list(df0M$line), mean)
colnames(df0M_mean)<-c("line","pupaMmean_0")
head(df0M_mean)

pupa_mean_M01<-merge(x=df0M_mean, y=df1M_mean, by.x="line", by.y="line")
head(pupa_mean_F01) #151

pupa_mean_M01$PM01<-pupa_mean_M01$pupaMmean_0-pupa_mean_M01$pupaMmean_1 #calculating SSP01
plastM01<-pupa_mean_M01 

```


```{r}
#SSP
pupa_SSP01<-merge(x=plastF01, y= plastM01, by.x="line", by.y="line") #merging male and female datasets
head(pupa_SSP01)
length(unique(pupa_SSP01$line)) #133 lines left

#calculate SSP from the plasticity.
pupa_SSP01$SSP01<-pupa_SSP01$PF01-pupa_SSP01$PM01

#calculating SSDs again to add them to the dataframe
pupa_SSP01$SSD0<-pupa_SSP01$pupaFmean_0-pupa_SSP01$pupaMmean_0
pupa_SSP01$SSD1<-pupa_SSP01$pupaFmean_1-pupa_SSP01$pupaMmean_1
head(pupa_SSP01,2)
```

#### Plot 
```{r}
SSP01_plot<-ggplot(pupa_SSP01, aes(x=reorder(line,SSP01), y=SSP01)) +
  geom_col(aes()) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  labs(x = " DGRP line ",y="SSP starved flies")
SSP01_plot

```

We saw that SSD variation in both food conditions is correlated with female size variation. Do we then have SSP variation due to female size variation or possibly plasticity in one of the two sexes?


## Question 3.3: which sex is more plastic?

```{r}
model1<-lmer(pupa~sex*day+(sex*day|line)+(1|block), data=df_sub01) #basic model
summary(model1) 

```

Look at the intercept values
```{r}
plasticityF = (14.566230) - (14.566230-0.179327)
print(c("plasticity female", plasticityF))
plasticityM = (14.566230 -0.094440 )- (14.566230-0.094440-0.179327+0.030618)
print(c("plasticity male", plasticityM))
```
Females are more plastic than males

Plot of female and male plasticity
```{r}
head(pupa_SSP01,2)

```






## Question 3.3: Is SSP genetic variation correlated with female size in either fed or starved conditions?

#### Correlation between SSP and female size in both conditions
Prepare a dataset with line, SSP, fed female size and starved female size

```{r}
head(df1_mean_F)
head(df0_mean_F)
head(pupa_SSP01)

merge1<-merge(x=df1_mean_F,y=df0_mean_F,by.x="line",by.y="line")
SSP_female01<-merge(x=merge1,y=pupa_SSP01,by.x="line",by.y="line")
head(SSP_female01)
SSP_female01<-SSP_female01[,c(1,8,9,14)]
SSP_female01
```
##### Plot
```{r}
#reshape data
SSP_female01<-gather(SSP_female01,key="day",value="pupaF",pupaFmean_0:pupaFmean_1)
#plot
sp4 <- ggscatter(SSP_female01, x = "SSP01", y = "pupaF",
   color = "day",
   add = "reg.line", conf.int = TRUE, alpha=0.5)
sp4 + ggpubr::stat_cor(aes(color = day)) #stat_cor uses the Pearson correlation
```

## Question 3.4: Is SSP genetic variation correlated male or female plasticity?

#### Correlation between SSP and plasticity
Prepare a dataset with line, SSP, fed female and male plasticity

```{r}
head(pupa_SSP01)
SSP_plast<-pupa_SSP01[,c(1,4,7,8)]
head(SSP_plast)

```

##### Plot
```{r}
#reshape data
SSP_plast<-gather(SSP_plast,key="sex",value="plasticity",PF01:PM01)
#plot
sp5 <- ggscatter(SSP_plast, x = "SSP01", y = "plasticity",
   color = "sex",
   add = "reg.line", conf.int = TRUE, alpha=0.5)
sp5 + ggpubr::stat_cor(aes(color = sex)) #stat_cor uses the Pearson correlation
```
There is a correlation between SSP and female plasticity.


























