---
title: "Statistics for DGRP data on SSD and SSP"
author: "Isabelle Vea"
date: "14/08/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
require(ggplot2)
require(car)
require(reshape2)
require(data.table)
require(plyr)
require(lme4)
require(lmerTest)
require(nlme)
require(stringr)
require(afex) #allows to retrieve p-values from lm4 function
require(stringr)
require(tidyr)
require(qqman)
require(tibble)
#devtools::install_github("variani/lme4qtl")
#require(lme4qtl)
require(lattice)
require(snow)
require(Rmpi)
require(ggfortify)
require(rapport)
require(pbapply)
require(svMisc)
```

```{r}
setwd("/Users/isabelle/Dropbox/_Github_reps/DGRP_SSDSSP/")

df<-read.csv("/Users/isabelle/Dropbox/_Github_reps/DGRP_SSDSSP/Sept2020/DGRPfinal_clean.csv")

head(df)
names(df)

```

```{r}
#changing lineage and day of starvation as character
df$Cline<-as.character(df$line)
df$Cday_of_starvation<-as.character(df$day_of_starvation)

typeof(df$line)
typeof(df$Cline)

```

```{r}
#subset experiment flies
df<-subset(df, df$control_experiment == 'EXPERIMENT')
head(df)
```


## Overall SSD regardless of treatment
```{r}
head(df)
t.test(df$lnpupa ~ df$sex)
f<-subset(df, df$sex == 'F')
m<-subset(df, df$sex == 'M')

summary(f$lnpupa) #mean 14.45
summary(m$lnpupa) #mean 14.38

```

## SSD by day starvation

```{r}
#SSD 0
SSD0<-subset(df, df$day_of_starvation ==0)
t.test(SSD0$lnpupa ~ SSD0$sex)

SSD1<-subset(df, df$day_of_starvation ==1)
t.test(SSD1$lnpupa ~ SSD1$sex)

SSD2<-subset(df, df$day_of_starvation ==2)
t.test(SSD2$lnpupa ~ SSD2$sex)
#In all day after starvation, females are larger than males on average

```




```{r}
#wing plot
#dfplot<- ggplot(df,aes(x=as.factor(Cday_of_starvation),y=log(df$wing)))+ 
# geom_boxplot(width=0.5, size=0.2, outlier.size = 0.1, aes(fill=sex)) + 
#  geom_jitter(aes(color=sex),position=position_jitterdodge(dodge.width=0.5,jitter.width = 0),size=0.5) +
#  theme_classic(base_size = 12) +
  #scale_fill_manual(name="", # Legend label, use darker colors
                  #breaks=c("Virgin methanol", "Mated","JHM"),
                  #labels=c("A","B", "C"),
                  #values=c("#2E9FDF","#FF0000","#E7B800")) +
#  scale_color_manual(values =c("black","black")) +
#  theme(strip.background=element_rect(fill="black"))+ 
#  theme(strip.text=element_text(color="white", face="bold"))+
#  facet_wrap(. ~ Cline)
#dfplot

```

#Anova

```{r}
names(df)
head(df)

```



#Testing SSD by sex and lineage

```{r}                      
model1<-lmer(lnpupa~sex+(1|line))


```



## Anova on lnwing by sex, day of starvation and lineage with interaction
```{r}
lm1<-lm(lnwing ~ sex * Cday_of_starvation * Cline, data = df)
Anova(lm1)

```

## Anova on lnleg by sex, day of starvation and lineage with interaction
```{r}
lm2<-lm(lnleg ~ sex * Cday_of_starvation * Cline, data = df)
Anova(lm2)

```


## Anova on lnwing by sex, day of starvation as integer and lineage with interaction
```{r}
lm1B<-lm(lnwing ~ sex * day_of_starvation * Cline, data = df)
Anova(lm1B)

```

##SSP day 0 and day 1

```{r}
df01<-subset(df, day_of_starvation <=1)
lm1d<-lm(lnwing ~ sex * Cday_of_starvation * Cline, data = df01)
Anova(lm1d)

lm1e<-lm(lnwing ~ sex * day_of_starvation * Cline, data = df01)
Anova(lm1e)
```


##Test for SSD

```{r}
df0<-subset(df,day_of_starvation == 0)

lm1c<-lm(lnwing ~ sex * Cline, data = df0)
Anova(lm1c)


lm2c<-lm(lnleg~ sex * Cline, data = df0)
Anova(lm2c)
```

If using day of starvation as an integer, the interaction is not significant



#subsetting by day 0 and Day 1 of starvation only

```{r}
unique(df$day_of_starvation)
df2<-subset(df, day_of_starvation < 2)
unique(df2$day_of_starvation)
unique(df2$Cday_of_starvation)
```

```{r}
lm4<-lm(lnwing ~ sex * Cday_of_starvation * Clineage, data = df2)
Anova(lm4)

```

```{r}
lm5<-lm(lnwing ~ sex * Cday_of_starvation * Clineage, data = df2)
Anova(lm5)

```


```{r}
df0<-subset(df,day_of_starvation ==0)
head(df0)
#wing plot
dfplot<- ggplot(df0,aes(x=sex,y=log(df0$wing)))+ 
  geom_boxplot(width=0.3, size=0.2, outlier.size = 0.1, aes(fill=sex)) + 
  geom_jitter(aes(color=sex),position=position_jitterdodge(dodge.width=0.5,jitter.width = 0),size=0.5) +
  theme_classic(base_size = 12) +
  #scale_fill_manual(name="", # Legend label, use darker colors
                  #breaks=c("Virgin methanol", "Mated","JHM"),
                  #labels=c("A","B", "C"),
                  #values=c("#2E9FDF","#FF0000","#E7B800")) +
  scale_color_manual(values =c("black","black")) +
  theme(strip.background=element_rect(fill="black"))+ 
  theme(strip.text=element_text(color="white", face="bold"))+
  facet_wrap(. ~ line)
dfplot

```

```{r}
df0<-subset(df,day_of_starvation ==0)
head(df0)
#wing plot
dfplot<- ggplot(df0,aes(x=sex,y=log(df0$leg)))+ 
  geom_boxplot(width=0.3, size=0.2, outlier.size = 0.1, aes(fill=sex)) + 
  geom_jitter(aes(color=sex),position=position_jitterdodge(dodge.width=0.5,jitter.width = 0),size=0.5) +
  theme_classic(base_size = 12) +
  scale_fill_manual(name="", # Legend label, use darker colors
                  breaks=c("M", "F"),
                  labels=c("M","F"),
                  values=c("#ef8a62","#67a9cf")) +
  scale_color_manual(values =c("black","black")) +
  theme(strip.background=element_rect(fill="black"))+ 
  theme(strip.text=element_text(color="white", face="bold"))+
  facet_wrap(. ~ line)
dfplot

```


SSD and SSP correlation

```{r}
SSPSSD<-read.csv(("SSP01_SSD_Aug032020.csv"))
head(SSPSSD)


library("ggpubr")
ggscatter(SSPSSD, x = "SSD_wing", y = "SSP_1_0_wing", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Wing Sexual Size Dimorphism",
          ylab = "Wing Sex Specific Plasticity")


```



## FDR for wing
```{R}
out<- read.csv("/Users/isabelle/Downloads/DGRP-June2020/July2020_data/DGRP2_outputs/wing_DGRP_malefem/ch_pos_mixedpval.txt", sep=' ')

head(out)

out$BH<-p.adjust(out$DiffMixedPval, method="BH")

out$BONF<-p.adjust(out$DiffMixedPval, method="bonferroni")

ordered_out <- out[order( out['BONF'] ),]
ordered_out <- out[order( out['BH'] ),]
ordered_out <- out[order( out['DiffMixedPval'] ),]


head(out)

```


#FDR by chromosome

```{r}
outwingSSD<- read.csv("/Users/isabelle/Downloads/DGRP-June2020/July2020_data/DGRP2_outputs/wing_DGRP_malefem/ch_pos_mixedpval.txt", sep=' ')
head(outwingSSD)
outwingSSD_2<-subset(outwingSSD, chromosome <=2 )
outwingSSD_3<-subset(outwingSSD, chromosome >=3 &&chromosome >=4)

outwingSSD_2$BH<-p.adjust(outwingSSD_2$DiffMixedPval, method="BH")
outwingSSD_2 <- outwingSSD_2[order( outwingSSD_2['BH'] ),]


```




## FDR for leg
```{R}
outSSD<- read.csv("/Users/isabelle/Downloads/DGRP-June2020/July2020_data/DGRP2_outputs/leg_DGRP2_malefem/ch_pos_DiffMixedPval.txt", sep=' ')
outSSP<- read.csv("/Users/isabelle/Downloads/DGRP-June2020/July2020_data/DGRP2_outputs/legplasticity-Aug2020/ch_pos_DiffMixedPval.txt", sep=' ')

head(outSSD)
head(outSSP)

outSSD$BH<-p.adjust(outSSD$DiffMixedPval, method="BH")
outSSD$BONF<-p.adjust(outSSD$DiffMixedPval, method="bonferroni")

ordered_outSSD_BONF <- outSSD[order( outSSD['BONF'] ),]
ordered_outSSD_BH <- outSSD[order( outSSD['BH'] ),]
ordered_outSSDpval <- outSSD[order( outSSD['DiffMixedPval'] ),]


outSSP$BH<-p.adjust(outSSP$DiffMixedPval, method="BH")
outSSP$BONF<-p.adjust(outSSP$DiffMixedPval, method="bonferroni")
ordered_outSSP_BONF <- outSSP[order( outSSP['BONF'] ),]
ordered_outSSP_BH <- outSSP[order( outSSP['BH'] ),]
ordered_outSSPpval <- outSSP[order( outSSP['DiffMixedPval'] ),]




head(out)

```


```{r}
head(SSPSSD)
plot(data=SSPSSD, male_1_0_wing~female_1_0_wing)
SSPSSD$exp_SSP_1_0_wing<-exp(SSPSSD$SSP_1_0_wing)
SSPSSD$exp_SSD_wing<-exp(SSPSSD$SSD_wing)

head(SSPSSD)


```




#How much of the environment is responsible for female size variation?
```{r}
#run Anova of size variation by DGRP line and extract the residuals
f<-subset(df, df$sex == 'F')
lmFline<-lm(lnpupa ~ Cline, data = f)
Anova(lmFline)
```

#How much of plasticity is responsible for female size variation?

```{r}
#I need Size plasticity, mean size in females

SSPSSD<-read.csv((""))

```